---
title: "area-under_curve"
author: "Stacey Harmer"
date: "7/20/2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

My goal is to automate a previously created function so that I can easily determine qPCR area-under-curve for individual genes and across unique time frames (time frames unique to each genotype and represent one circadian day).

```{r}
library(tidyverse)
library(MESS)
```

Below is the original code:

```{r}
Akiva.12deg<- read.csv("qRT-PCR_data_12c_6genes_n.csv")
head(Akiva.12deg)
tail(Akiva.12deg)

```

Make the data wide
```{r}
PCR.12.wide <- spread(Akiva.12deg,key = gene, value = expn ) %>%
  mutate(gt.rep = paste(gt, rep, sep = "."))
head(PCR.12.wide)

```

Original code:

```{r}

PRR5_12deg_24_48.6 <- sapply(unique(PCR.12.wide$gt.rep), function(gt.rep) {
  tmp <- PCR.12.wide[PCR.12.wide$gt.rep==gt.rep,]
  auc(tmp$time,tmp$PRR5,type="spline", from =24, to = 48.6)
})
PRR5_12deg_24_48.6
```

I want to be able to step through each of the genes.  would it be better if I used long insetad of wide dataframe?

```{r}

Akiva.12deg <- Akiva.12deg %>%
  mutate(gt.rep = paste(gt, rep, sep = "."))

head(Akiva.12deg)

#all_12deg_24_48.6 <- sapply(unique(Akiva.12deg$gt.rep), function(gt.rep) {
#  gene <- Akiva.12deg[Akiva.12deg$gene==gene,],
#  tmp <- Akiva.12deg[Akiva.12deg$gt.rep==gt.rep,]
#  auc(tmp$time,tmp$gene,type="spline", from =24, to = 48.6)
#})

```
that failed. 

```{r}
head(PCR.12.wide)
```

I suspect I need to incoroprate something like the below

 for(i in 5:10)
    { 
      PCR.12.wide[, i];
    }
but I can't figure it out

