---
title: "area-under_curve"
author: "Stacey Harmer"
date: "7/20/2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

My goal is to automate a previously created function so that I can easily determine qPCR area-under-curve for individual genes and across unique time frames (time frames unique to each genotype and represent one circadian day).

```{r}
library(tidyverse)
library(MESS)
library(stringr)
```

Below is the original code:

```{r}
Akiva.12deg<- read_csv("qRT-PCR_data_12c_6genes_n.csv") #read it in as a tibble
head(Akiva.12deg)
tail(Akiva.12deg)

```

Make the data wide
```{r}
PCR.12.wide <- spread(Akiva.12deg,key = gene, value = expn ) %>%
  mutate(gt.rep = paste(gt, rep, sep = "."))
head(PCR.12.wide)

```

Original code:

```{r}

PRR5_12deg_24_48.6 <- sapply(unique(PCR.12.wide$gt.rep), function(gt.rep) {
  tmp <- PCR.12.wide[PCR.12.wide$gt.rep==gt.rep,]
  auc(tmp$time,tmp$PRR5,type="spline", from =24, to = 48.6)
})
PRR5_12deg_24_48.6
```

I want to be able to step through each of the genes.  would it be better if I used long insetad of wide dataframe?

```{r}

Akiva.12deg <- Akiva.12deg %>%
  mutate(gt.rep = paste(gt, rep, sep = "."))

head(Akiva.12deg)

#all_12deg_24_48.6 <- sapply(unique(Akiva.12deg$gt.rep), function(gt.rep) {
#  gene <- Akiva.12deg[Akiva.12deg$gene==gene,],
#  tmp <- Akiva.12deg[Akiva.12deg$gt.rep==gt.rep,]
#  auc(tmp$time,tmp$gene,type="spline", from =24, to = 48.6)
#})

```
that failed. 

```{r}
head(PCR.12.wide)
```

I suspect I need to incoroprate something like the below

 for(i in 5:10)
    { 
      PCR.12.wide[, i];
    }
but I can't figure it out

Yes, that would be one way to do it, but I wonder if we could use summarize instead

```{r}
results <- Akiva.12deg %>%
  group_by(gene,gt,rep) %>%
  summarize(auc=auc(time,expn,type="spline", from =24, to = 48.6))
```

let's compare the prr5 results from this method and the Stacey's single call
```{r}
results %>% filter(gene=="PRR5")
```

```{r}
PRR5_12deg_24_48.6
```

Success!

### can we use lmer as function within summarize?

I will calculate the AUC for one circadian day for the differnet genotypes

### Col 
```{r}
library(MESS)

auc_Col <- Akiva.12deg %>%
  group_by(gene, rep, gt) %>%
  summarize(auc = auc(time, expn, type="spline", from =24, to = 48.6)) %>%
  filter(gt == "Col") %>%
  mutate(gt.biorep = paste(gt, rep, sep="."))

auc_Col
# and now spread so it's wide 

auc_Col <- spread(auc_Col, key = gene, value = auc )
```

### cca1 lhy
ccalhy: 21.5 hr   (24 to 45.5)
```{r}
auc_cca1lhy <- Akiva.12deg %>%
  group_by(gene, rep, gt) %>%
  summarize(auc = auc(time, expn, type="spline", from =24, to = 45.5)) %>%
  filter(gt == "cca1lhy") %>%
  mutate(gt.biorep = paste(gt, rep, sep="."))

auc_cca1lhy
# and now spread so it's wide 

auc_cca1lhy <- spread(auc_cca1lhy, key = gene, value = auc )

```

### rve468
rve468: 29.24     (24 to 53.2)
```{r}
auc_rve468 <- Akiva.12deg %>%
  group_by(gene, rep, gt) %>%
  summarize(auc = auc(time, expn, type="spline", from =24, to = 53.2)) %>%
  filter(gt == "rve468") %>%
  mutate(gt.biorep = paste(gt, rep, sep="."))

auc_rve468
# and now spread so it's wide 

auc_rve468 <- spread(auc_rve468, key = gene, value = auc )

```

### quint
quint: 25.5      (24 to 49.5)
```{r}
auc_quint <- Akiva.12deg %>%
  group_by(gene, rep, gt) %>%
  summarize(auc = auc(time, expn, type="spline", from =24, to = 49.5)) %>%
  filter(gt == "cca1lhyrve468") %>%
  mutate(gt.biorep = paste(gt, rep, sep="."))

auc_quint
# and now spread so it's wide 

auc_quint <- spread(auc_quint, key = gene, value = auc )

```

Great, now join them together.  

```{r}

auc_all <- bind_rows(auc_Col, auc_cca1lhy, auc_rve468, auc_quint)
auc_all
```

Now gather for easy ggplotting
```{r}
auc_all_gather <- auc_all %>%
  gather(key = "gene", value ="AUC", 4:9)

```

Now I can easily plot them all using facet

```{r}
# first, order genotypes appropriately
gt_order <- c("Col", "cca1lhy", "rve468", "cca1lhyrve468")

auc_all_gather <- auc_all_gather %>%
  mutate(gt = factor(gt, levels = gt_order)) %>%
  arrange(gt)

# now plot
auc_all_gather %>%
  group_by(gt) %>%
  ggplot(aes(color = gt)) +
  geom_boxplot(aes(x = gt, y=AUC)) + 
  facet_wrap(~gene)

```

Can I use lmer wiht summrize?

```{r}
library(lme4)
library(lmerTest)
head(auc_all_gather)
```

My attempt:

Rather than comment this out, use eval=FALSE in the header.  Makes it easier to run and see the error when you want to

```{r, eval=FALSE}

lmer_12deg <-  auc_all_gather %>%
  group_by(gene, gt) %>%
  summarize(lmer = lmer(AUC ~ gt + (1|rep))) 

```

failed.

First, although this is not causing your error, I don't think you want to be grouping by `gt` since that is in your `lmer` formula.  If you do, then you are fitting a separate lmer model both for every gene and for every genotype.

The first problem is that `%>%` feeds in the tibble to the first argument of the function (`lmer` in this case) but `lmer` doesn't want it there.  So we have to explicitly define where we want the tibble to go with `.`

```{r, eval=FALSE}

lmer_12deg <-  auc_all_gather %>%
  group_by(gene) %>%
  summarize(lmer = lmer(AUC ~ gt + (1|rep), data = .)) 

```

But in this case it is actually worse because summarize expects us to name which columns we want to work on.

An additional problem is that summarize wants the function to produce a single value, not a complicated object.  So you could do something liket this, but it is ugly.  Plus lmer is getting called 7 times for every gene instead of just once.

_If you are going to use this, please double check that it is giving the expected results...

```{r}
lmer.summary <- function(AUC,gt,rep,coef,column) {
     tmp <- summary(lmer(AUC ~ gt + (1|rep)))[["coefficients"]][coef,column] 
     return(tmp)
}

lmer_12deg <-  auc_all_gather %>%
 group_by(gene) %>%
 summarize(Col.estimate = lmer.summary(AUC,gt,rep,coef="(Intercept)",column="Estimate"),
           cca1lhy.estimate = lmer.summary(AUC,gt,rep,coef="gtcca1lhy",column="Estimate"),
           cca1lhy.pvalue = lmer.summary(AUC,gt,rep,coef="gtcca1lhy",column="Pr(>|t|)"),
           rve468.estimate = lmer.summary(AUC,gt,rep,coef="gtrve468",column="Estimate"),
           rve468.pvalue = lmer.summary(AUC,gt,rep,coef="gtrve468",column="Pr(>|t|)"),
           cca1lhyrve468.estimate = lmer.summary(AUC,gt,rep,coef="gtcca1lhyrve468",column="Estimate"),
           cca1lhyrve468y.pvalue = lmer.summary(AUC,gt,rep,coef="gtcca1lhyrve468",column="Pr(>|t|)")
           )
lmer_12deg
```

Trying an alternative way to specify the arguments to summarize...would still run lmer 7X more than needed.  And the setup is ugly...
And doesn't work
```{r, eval=FALSE}
names <- str_c(rep(gt_order,each=2),c("estimate","pvalue"),sep=".")
coefs <- rep(c("(Intercept)",str_c("gt",unique(auc_all_gather$gt))),each=2)
dots <- str_c("lmer.summary(AUC,gt,rep,coef='",coefs,"',column='",c("Estimate","Pr(>|t|)"),"')")

lmer_12deg <-  auc_all_gather %>%
 group_by(gene) %>%
 summarize_(.dots=setNames(dots,names)
           )
lmer_12deg

```


Instead we should use lapply (I think we will learn some tidy alternatice ttechniques for this in a few chapters)

```{r}
lmer_12deg <- lapply(unique(auc_all_gather$gene) , function(gene) {
    lmer(AUC ~ gt + (1|rep), data=auc_all_gather[auc_all_gather$gene == gene,])
})
names(lmer_12deg) <- unique(auc_all_gather$gene)
```

Now we have a list of lmer objects and we can do various operatios on them

```{r}
lmer.summaries <- lapply(lmer_12deg,summary)
```

```{r}
lmer.coefficient.tables <- lapply(lmer.summaries,function(x) x[["coefficients"]])
lmer.coefficient.tables
```

```{r}
combined.coefficient.table <- plyr::ldply(lmer.coefficient.tables)
#unfortunately row names are lost
combined.coefficient.table$gt <- unlist(lapply(lmer.coefficient.tables,rownames))
combined.coefficient.table
```

